---
title: "Fit a skew-normal distribution to DNA fraction deistribution"
author: "Roey Angel"
email: "roey.angel@bc.cas.cz"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_document:
    code_folding: show
    dev: png
    df_print: kable
    fig_caption: yes
    highlight: pygments
    keep_md: yes
    number_sections: no
    theme: flatly
    toc: yes
    toc_depth: 5
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '5'
link-citations: yes
csl: fems-microbiology-ecology.csl
subtitle: ""
---

```{r libraries, include=F}
library(tidyverse) # Easily Install and Load the 'Tidyverse', CRAN v1.3.1
library(extrafont) # Tools for using fonts, CRAN v0.17
library(kableExtra) # Construct Complex Table with 'kable' and Pipe Syntax, CRAN v1.3.4
library(knitr) # A General-Purpose Package for Dynamic Report Generation in R, CRAN v1.33
library(magrittr) # A Forward-Pipe Operator for R, CRAN v2.0.1
library(sn) # The Skew-Normal and Related Distributions Such as the Skew-t and the SUN, CRAN v2.0.0
library(patchwork) # The Composer of Plots, CRAN v1.1.1
library(svglite) # An 'SVG' Graphics Device, CRAN v2.0.0
```

```{r style settings, include=F}
options(width = 90, knitr.table.format = "html")
opts_chunk$set(
  warning = FALSE,
  message = FALSE,
  cache = TRUE,
  dev = "png",
  fig.ext = "png",
  dpi = 600,
#  fig.width = 12,
#  fig.height = 8,
  cache.path = "Fit_selm_cache/",
  fig.path = "Fit_selm_figs/"
)
f_name <- "DejaVu Sans" #sub("\\s//", "", f_name)
f_size <- 10
font_import(pattern = "DejaVuSans", prompt = FALSE)
loadfonts() # registers fonts
theme_set(theme_bw(base_size = f_size, base_family = f_name))
```

## Functions
```{r functions, include=T}
bp2kbp <- function() {
    function(x) format(x/1000,digits = 2) 
}

plot_fitted_dist <- function(frag_sizes = DNA_frags, mode_sample = rsamp1){
  ggplot() +
    geom_point(data = frag_sizes, aes(x = `Fragment size (bp)`, y = Intensity, colour = "Data")) +
    geom_freqpoly(data = mode_sample, aes(x = sn, colour = "Model fit SN")) +
    # geom_freqpoly(data = rsamp2, aes(x = st, colour = "Model fit ST")) +
    # geom_freqpoly(data = rsamp3, aes(x = sn, colour = "Model fit SN log")) +
    # geom_freqpoly(data = rsamp4, aes(x = st, colour = "Model fit ST log")) +
    # geom_freqpoly(data = rsamp_guess, aes(x = gs, colour = "Guess")) +
    scale_x_log10(limits = c(300, 60000), labels = bp2kbp()) +
    scale_colour_manual(name = "", breaks = c("Data","Model fit SN", "Model fit ST", "Model fit SN log", "Model fit ST log", "Guess"),
                        values = c("Data" = "black",
                                   "Model fit SN" = "red",
                                   "Model fit SN log" = "blue",
                                   "Model fit ST" = "green",
                                   "Model fit ST log" = "Orange",
                                   "Guess" = "gray")) +
    theme(legend.position = c(0.15, 0.95), 
          plot.title = element_text(color = "black", size = 12, hjust = 0.5, face = "bold"),
          legend.background = element_blank()) +
    labs(x = "Fragment size (kbp)")
    
    
     # labs(colour = "")
    # theme(legend.position = "none")
}

# convert svg to svgz
gz <- function (in_path, out_path = tempfile()) 
{
    out <- gzfile(out_path, "w")
    writeLines(readLines(in_path), out)
    close(out)
    file.remove(in_path)
    invisible(out_path)
}
```

# Model the distribution parameters of DNA fragments from environmental samples using the skew-normal distribution

## Set DNA fraction sizes
```{r set general parameters, cashe=T}
rel_path <- "../Agilent_tape_station/2020-07-13 - 18-32-37_Angel-gDNA/"

# Set fragment sizes manually according to selected regions because GetData Graph Digitizer cannot handle a log x-axis well
read_csv(paste0(rel_path, "2020-07-13 - 18-32-37_Angel-gDNA_regionTable.csv"), 
         skip = 2,
         col_names = FALSE, 
         n_max = 31) %>% 
  select(X1, X2) %>% 
  pivot_longer(X1:X2, values_to = "Fragment size (bp)") %>% 
  pull(`Fragment size (bp)`) -> 
  Fragment_sizes

dist_params <- tibble(Sample = character(),
                      mean = numeric(),
                      s.d. = numeric(),
                      gamma1 = numeric(),
                      xi = numeric(),
                      omega = numeric(),
                      alpha = numeric())


N <- 10000 # n points to sample for plotting
```

## Temperate forest soil I (Bohemian Forest)
```{r B1 Sumava 1, cache=T}
# this_max <- 6950 # maximum of peak (if recorded)
# Manual_fragment_sizes_many %>% append(., this_max, after = sum(this_max > .)) -> Fragment_sizes
Sample_name <- "Bohemian Forest soil I"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_B1_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)
# DNA_frags %<>% bind_cols(., `Log fragment size (bp)` =  log10(DNA_frags$`Fragment size (bp)`))

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, ., .id = NULL) ->
  dist_params
confint(fit1)
summary(fit1)

# fit2 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "ST", method = "MPLE")
# print(fit2@param)
# 
# fit3 <- selm(`Log fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
# plot(fit3, which = 4)
# print(fit3@param)
# 
# fit4 <- selm(`Log fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "ST")
# print(fit4@param)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 1.05))

# rsamp2 <- data.frame(st = rsn(fit2, dp = coef(fit2, "DP"), 
#                               n = N/3)) # ignoring nu because I cannot use it for scipy.skewnorm
# 
# # fitted data based on log model
# rsamp3 <- data.frame(sn = 10^(rsn(fit3, dp = coef(fit3, "DP")[1:3],
#                               n = N/2)))
# 
# rsamp4 <- data.frame(st = 10^(rsn(fit4, dp = coef(fit4, "DP")[1:3],
#                               n = N/3)))
# 
# rsamp5 <- data.frame(st = rsn(xi = coef(fit1, "DP")[1],
#                               omega = coef(fit1, "DP")[2],
#                               alpha = -0.4,
#                               n = N))
# 
# 
# just guessing the params
# rsamp_guess <- data.frame(gs = rsn(dp = cp2dp(c(6950, 5000, 0.7), "SN"),
# n = N))

p1 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Temperate forest soil I \n Bead beating + phenol") + theme(plot.title = element_text(size = 12))
print(p1)
```

## Temperate forest soil II (Bohemian Forest)
```{r C1 Sumava 2, cache=T}
# this_max <- 7428 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "Bohemian Forest soil II"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_C1_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N * 1.05))

p2 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Temperate forest soil II \n Bead beating + phenol") + theme(plot.title = element_text(size = 12)) 
print(p2)
```

## Vienna city soil
```{r H2 City soil, cache=T}
# this_max <- 6155 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "Vienna city soil"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_H2_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)
# DNA_frags %<>% bind_cols(., `Log fragment size (bp)` =  log10(DNA_frags$`Fragment size (bp)`))

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 10.5))
p3 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Vienna garden soil \n DNeasy PowerSoil Kit") + theme(plot.title = element_text(size = 12)) 
print(p3)
```

## Austrian Gleysol
```{r G2 Teabag soil, cache=T}
# this_max <- 7852 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "Austrian Gleysol"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_G2_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)
# DNA_frags %<>% bind_cols(., `Log fragment size (bp)` =  log10(DNA_frags$`Fragment size (bp)`))

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = `Intensity`, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 2.3))
p4 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Austrian Gleysol \n DNeasy PowerSoil Kit") + theme(plot.title = element_text(size = 12)) 
print(p4)
```

## Arizona desert soil
```{r H1 Desert Baboquisari, cache=T}
# this_max <- 6803 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "Arizona desert soil"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_H1_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)
# DNA_frags %<>% bind_cols(., `Log fragment size (bp)` =  log10(DNA_frags$`Fragment size (bp)`))

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 5.2))
p5 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Arizona desert soil \n DNeasy PowerSoil Kit?") + theme(plot.title = element_text(size = 12)) 
print(p5)
```

## Arctic permafrost soil
```{r D4 Arctic soil, cache=T}
# this_max <- 7641 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "Arctic permafrost soil"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_D4_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)
# DNA_frags %<>% bind_cols(., `Log fragment size (bp)` =  log10(DNA_frags$`Fragment size (bp)`))

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 3.3))
p6 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Arctic permafrost soil \n ?") + theme(plot.title = element_text(size = 12)) 
print(p6)
```

## Post-mining reclamation lake periphyton
```{r B3 Post-mining reclamation lake, cache=T}
# this_max <- 7920 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "Post-mining reclamation lake periphyton"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_B3_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
# confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 4.5))
p7 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Lake periphyton \n DNeasy PowerSoil Kit") + theme(plot.title = element_text(size = 12))
print(p7)
```

## Drylands millipede hindgut (*A. syriacus*)
```{r A2 Millipede (tropical hindgut), cache=T}
# this_max <- 8218 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <-"Drylands millipede hindgut"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_A2_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N / 1.6))

p8 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "Tropical millipede \n FastDNA Spin Kit for Soil") + theme(plot.title = element_text(size = 12)) 
print(p8)
```

## European millipede hindgut (*G. connexa*)
```{r B2 Millipede (European hindgut), cache=T}
# this_max <- 9447 # maximum of peak (if recorded)
# Fragment_sizes %<>% append(., this_max, after = sum(this_max > .))
Sample_name <- "European millipede hindgut"
read_table(paste0(rel_path, "2020-07-13 - 18.32.37_Angel_B2_regions.txt"),
           col_names = c("Fragment size (bp)", "Intensity"), 
           skip = 4) %>% 
  mutate(Intensity = round(Intensity)) -> # weights in selm must be integers
  DNA_frags

DNA_frags$`Fragment size (bp)` <- Fragment_sizes 
DNA_frags %>% glimpse

DNA_frags %<>% filter(., `Fragment size (bp)` > 350)
DNA_frags %<>% filter(., `Fragment size (bp)` < 50000)

fit1 <- selm(`Fragment size (bp)` ~ 1, weights = Intensity, data = DNA_frags, family = "SN")
plot(fit1, which = 4)
print(fit1@param)
fit1@param[2:1] %>% 
  flatten() %>% 
  enframe() %>% 
  unnest(cols = c(value)) %>% 
  spread(key = name, value = value) %>% 
  bind_cols(Sample = Sample_name, .) %>% 
  bind_rows(dist_params, .) ->
  dist_params
confint(fit1)
summary(fit1)

# fitted data based on models
rsamp1 <- data.frame(sn = rsn(fit1, dp = coef(fit1, "DP"),
                                 n = N * 1.7))

p9 <- plot_fitted_dist(frag_sizes = DNA_frags, mode_sample = rsamp1) + labs(title = "European millipede \n FastDNA Spin Kit for Soil") + theme(plot.title = element_text(size = 12)) 
print(p9)
```

## Summary
```{r Summary, fig.show = 'hide', cache=T, fig.width=12, fig.height=14, out.width = '1000px'}

colnames(dist_params) <- c("Sample", "Mean", "SD", "Gamma", "Xi", "Omega", "Alpha")
dist_params %>% 
  kable(., digits = c(1, 1, 4)) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F)

dist_params %>% 
  filter(., Sample != "Post-mining reclamation lake periphyton") %>% # outlier
  select(-Sample) %>% 
  summarise(Mean = mean(Mean), SD = sqrt(mean(SD^2)), Gamma = mean(Gamma),
            Xi = mean(Xi), Omega = sqrt(mean(Omega^2)), Alpha = mean(Alpha)) %>% 
  kable(., digits = c(1, 1, 4), caption = "Mean distribution") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"), 
                full_width = F)
  
p.file <- "Distribution_fit_summary"
dist_fit_sum <- p1 + p2 + theme(legend.position = "none") + p3 + theme(legend.position = "none") + p4 + theme(legend.position = "none") + p5 + theme(legend.position = "none") + p6 + theme(legend.position = "none") + p7 + theme(legend.position = "none") + p8 + theme(legend.position = "none") + p9 + theme(legend.position = "none") +  plot_layout(ncol = 3)

dist_fit_sum 
svglite(paste0("Fit_selm_figs/", p.file, ".svg"), width = 12, height = 14)
print(dist_fit_sum )
dev.off()

gz(paste0("Fit_selm_figs/", p.file, ".svg"), paste0("Fit_selm_figs/", p.file, ".svgz"))
```

```{r colophon, eval=T}
sessioninfo::session_info() %>%
  details::details(
    summary = 'Current session info',
    open    = TRUE
 )
```